---
title: "5513 ASSIGNMENT 2"
author: "Pranali Angne"
Date: "`r Sys.Date()`"
output:
 bookdown::html_document2  
keep_md: yes
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

# Introduction

The Gun control issue is a growing concern in the States and the objective of this report is to explain the research question – Gun control, less crime? with the help of the following data set https://vincentarelbundock.github.io/Rdatasets/datasets.html that has been extracted from an open data source. 

In the data set below there are **5 variables** and **999 observations**. This report will compare these variables that will be shown in the report through data visualization. And analysis will bring the report to conclusion with the help of graphs, plots and tables. 

This analysis is based on the data  of United States of America during 22 years from 1977 to 1999. The objective of this report is to understand the trend, whether the crime since 1977 till 1999 has reduced or not and understanding which state has the highest number of murders and violence and is there any law or not to punish the guilty. 




```{r Chunk 1, message = FALSE, warning = FALSE, eval = TRUE}
library(dplyr)
library(ggplot2)
library(kableExtra)
```


```{r Reading data, echo = TRUE, warning = FALSE, message = FALSE}
library(readxl)
Gun <- read_excel("~/Desktop/DATA/data/Gun.xlsx")
dim(Gun)
```

The number of rows are `r {nrow(Gun)}` and the number of variables are `r {ncol(Gun)}` 

## Summary

```{r}
Gun_new <- c ("year", "murder", "state", "law", "violent")
summary(Gun [, Gun_new])
```


## Variable names

```{r}
names(Gun)
```

In the data set below there are 5 variables viz. 

**Year**, **Violent**, **Murder**, **State**, **Law**


# Data Analysis

**GRAPH: 1**

```{r Figref1,fig.cap = "Yearly murder analysis"}
plot1 <- Gun %>% dplyr::group_by(year) %>%
  count(murder)
ggplot(plot1, aes(y = murder  , x = year, fill = murder)) + geom_point() +
  theme_gray() 
```



## Results and discussion: 

The above graph \@ref(fig:Figref1) explains the year from 1977 till 1999. It tries to analyze that, has the murder trend changed since 1977 and what is the highest murder rate since then. So, the initial year from 1977 till 1985 the murders ranged somewhere around 20 to 35 murders. After 1985, the murder rate increased at an increasing rate and in the year 1991 the murder rate was the highest. After 1991 there is a decline in murders.


**GRAPH: 2**

```{r Figref2,fig.cap = "Yearly violence analysis" }
plot2 <- Gun %>% dplyr::group_by(year) %>%
  count(violent)
ggplot(plot2, aes(y = violent  , x = year, fill = violent)) + geom_point() +
  theme_gray() 
```

## Results and Discussion:
The above graph \@ref(fig:Figref2) explains, this analysis is similar to that of yearly murder analysis but here we see that procuring guns leads to more violence rather than death/murder. The year 1992 and 1993 records the highest number of violence occurred, but again there is a fall after a certain point of time. 


# Tables 

**TABLE: 1**

```{r tabref1}
table1 <- Gun %>% 
  dplyr::filter(year ==1977) %>% 
  select(state,violent, murder, law)
  
knitr::kable(head(table1, n=44), caption = "These are the first 64 rows") %>% 
  kable_styling(bootstrap_options = c("hover", "striped"))
```

## Result and discussion:

The table \@ref(tab:tabref1) shows **District of Columbia** has the highest number of violence along with highest number of murders too. Also, states like **New York**, **California**, **Maryland**, **Illinois**, and **Florida** are the top six states having the highest rate of murders and violence. Considering the year 1977, states having laws against Guns are very few and the top six states showing the highest rate of murders and violence have no law against Gun. 





**TABLE: 2**

```{r tabref2}
table2 <- Gun %>% dplyr::filter(year == 1999) %>% 
  select(state, violent, murder, law)

knitr::kable(head(table2, n=43), caption = "These are the first 64 rows") %>% 
  kable_styling(bootstrap_options = c("hover", "striped"))


```

## Result and discussion

Now looking at the table \@ref(tab:tabref2) of 1999, **District of Colombia** has still the highest number of violence and no law has been introduced against the Gun control. 
The only states out of the top six are  **New York** and **Florida** that has introduced a law against gun. Also, there are few more states progresses but yet the situation is not under control.  



# Conclusion for analysis

In short, gun ownership does not increase safety, and the prevalence of guns directly correlates with significantly greater risk of gun-related homicides and suicides. While the facts surrounding the safety of having a gun in the home are clear, the choice to own a gun is more complicated for many homeowners.

